Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-03-24 2_SSL_EES/article/article_supplementary.tex
updated figures

\begin{myequation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{D}  \frac{1}{y_i}   \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x \c_x \bigg\}
\end{myequation}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=5.5cm]{./figures/die-per-wafer.pdf}
    \caption{Number of die per wafer for the different wafer sizes used in different years: (DPW) $d(2002)=50$mm$\rightarrow851$DPW, $d(2002)=50$mm$\rightarrow5370$DPW, $d(2020)=100$mm$, \rightarrow26,838$DPW. Source: own estimates, based on \cref{eqn:dpw} and the dominant wafer size of the considered year, according to \cref{fig:wafers}.}
    \label{fig:dpw}
\end{figure}

Results are plotted in \cref{fig:dpw}, showing the increase in die per wafer (DPW) over time. \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category
\begin{myequation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x \c_x \bigg\}
\end{split}
\end{myequation}
and the cost of a manufacturing step $C_j$ in the packaging category
\begin{myequation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x c_x \bigg\}
\end{myequation}
where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{myequation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{D} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{myequation}
Note that in keeping with the categorization introduced by the United States Department of Energy (cf. eg. \cite{doe_ssl_rnd_2019}), certain steps from these two categories are reported separately. In the wafer processing category, the epitaxy step is reported separately due to its complexity and the large share of cost carried. In the wafer processing category, the phosphor step is reported separately.

@@

\begin{myequation}
    f_{\times 2} =  \frac{\frac{i+s}{y}+s}{y}
\end{myequation}
\begin{myequation}
    f_{\times n} = \frac{i + s(1+y+y^2+ \dots + y^{n-1})}{y^n}
\label{eqn:series}
\end{myequation}

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/costmodel_calibration.pdf}
	\caption{Modelled manufacturing cost for light-emitting diode packages split by manufacturing step category. The top graph shows the data from LEDCOM model and round table discussions published by the United States Department of Energy (DoE). The bottom graph shows the results of the cost model developed in this thesis. Hatched bars are projections given by the sources. Inset plot shows the number of die per wafer (DPW) calculated for different wafer sizes used in the model. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in Section \cref{sec:costmodel}. Sources for Inset Plot: own elaborations based on numerical approximations for die-per-wafer estimations provided in \cite{de2005investigation}, informed by expert interviews.}
	\label{fig:costmodel_calibration}
\end{figure}

\cref{fig:costmodel_calibration} shows a comparison of the relative manufacturing cost by manufacturing step categories between the cost model and data released by the United States Department of Energy (DoE). Note that any comparison between the s must consider: Our model assumes state-of-the-art equipment, while the DoE industry round table reports data from factories which use older equipment, as well as newer equipment. Our model presently considers only low-power and mid-power packages, while the DoE reports consider high-power packages also. Our model assumes a virtual North American manufacturing location, while the majority of manufacturers produce in Asia. We note that the share of the substrate price in the DoE data is much larger than in our model. This can in part be explained by  \cref{fig:sapphire_prices}, which already shows how projections overestimated the actual price for 2015-2020. We note also that the share of the epitaxy step is larger in the DoE model. This can in part be explained by our model relying on state-of-the-art equipment, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In conclusion, our model fits industry forecasts due to the way it is set up. The packaging part of the manufacturing process is increasing in importance, which has been confirmed by researchers and industry reports on WLP\cite{Lee2011WPL}\cite{Xie2013}\cite{ledsmag2017WLP}.

>>>>

\begin{myequation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{D}  \frac{1}{y_i}   \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x c_x \bigg\}
\end{myequation}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=5.5cm]{./figures/die-per-wafer.pdf}
    \caption{Number of die per wafer for the different wafer sizes used in different years: (DPW) $d(2002)=50$mm$\rightarrow851$DPW, $d(2002)=50$mm$\rightarrow5370$DPW, $d(2020)=100$mm$, \rightarrow26,838$DPW. Source: own estimates, based on \cref{eqn:dpw} and the dominant wafer size of the considered year, informed by expert interviews. For detailed statistics on changes in wafer size, compare also \cref{fig:wafers}.}
    \label{fig:dpw}
\end{figure}

Results are plotted in \cref{fig:dpw}, showing the increase in die per wafer (DPW) over time. \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category
\begin{myequation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x c_x \bigg\}
\end{split}
\end{myequation}
and the cost of a manufacturing step $C_j$ in the packaging category
\begin{myequation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((c_e*p) + c_l + c_m + c_d + c_o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x c_x \bigg\}
\end{myequation}
where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{myequation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{D} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{myequation}
Note that in keeping with the categorization introduced by the United States Department of Energy (cf. eg. \cite{doe_ssl_rnd_2019}), certain steps from these two categories are reported separately. In the wafer processing category, the epitaxy step is reported separately due to its complexity and the large share of cost carried. In the wafer processing category, the phosphor step is reported separately.

@@

\begin{myequation}
    f_{\times 2} =  \frac{\frac{i+s}{y}+s}{y}
\end{myequation}
\begin{myequation}
\label{eqn:series}
    f_{\times n} = \frac{i + s(1+y+y^2+ \dots + y^{n-1})}{y^n}
\end{myequation}

@@

\begin{figure}[h]
	\centering
    \includegraphics[width=\textwidth]{2_SSL_EES/article/figures/costmodel_calibration.pdf}
	\caption{Comparison of LED manufacturing cost structure between our manufacturing cost model (bottom panel) and previously published US DOE manufacturing cost calculations and projections based on industry data and LEDCOM model (top panel). Hatched bars are projections given by US DOE for 2018 and 2020. Sources for top panel: \cite{doe2010solid}\cite{doe2011solid}\cite{doe2012solid}\cite{doe2013solid}\cite{doe2014solid}\cite{doe2015solid}\cite{doe2016solid}. Sources for bottom panel: own elaboration based on the cost model described in Section \cref{sec:costmodel}.}
	\label{fig:costmodel_calibration}
\end{figure}
